---
title: "Tornado-Level Estimates using Areal"
author: "Tyler Fricker"
date: "1/14/2019"
output: html_document
---

Set working directory and load packages.
```{r}
library(tidyverse)
library(lubridate)
library(sf)
library(tmap)
library(USAboundaries)
library(rgeos)
```

### Data and Methods

Tornado data:

Download the tornado data from the Storm Prediction Center (SPC) http://www.spc.noaa.gov/gis/svrgis/ and load the shapefile into R.
```{r}
download.file(url = "http://www.spc.noaa.gov/gis/svrgis/zipped/1950-2017-torn-aspath.zip",
              destfile = "tornado2017.zip")
unzip("tornado2017.zip")
Torn.sf <- read_sf(dsn = "1950-2017-torn-aspath")
```

Remove tornadoes occurring in Hawaii, Alaska, and Puerto Rico and those occurring before 1994. That year marks the beginning of comprehensive WSR-88D radar coverage. For missing EF ratings use the modification rules (if/else) defined here: https://www.spc.noaa.gov/wcm/OneTor_F-scale-modifications.pdf
```{r}
Torn.sf <- Torn.sf %>%
  filter(yr >= 1995,
         !st %in% c("AK", "PR", "HI")) %>%
  mutate(mag = ifelse(mag == -9 & len <= 5, 0, mag),
         mag = ifelse(mag == -9 & len > 5, 1, mag))
```

Add a data/time column also add columns for path length, width, and area in metric units. Leave the time zone as native CDT.
```{r}
Torn.sf <- Torn.sf %>%
  mutate(dy = format(as.Date(date, format="%Y-%m-%d"), "%d"),
         DateTime = as.POSIXct(paste(yr, mo, dy, time), format = "%Y%m%d%H:%M:%S"),
         Hour = hour(DateTime),
         Year = year(DateTime),
         Length = len * 1609.34,
         Length = ifelse(Length == 0, min(Length[Length > 0]), Length), #takes care of zero length
         Width = wid * .9144,
         Width = ifelse(Width == 0, min(Width[Width > 0]), Width), #takes care of zero width
         Width = ifelse(Year >= 1995, Width * pi/4, Width), #takes care of change: avg to max
         cas = inj + fat,
         AreaPath = Length * Width,
         Ma = factor(month.abb[mo], levels = month.abb[1:12])) %>%
         filter(cas > 0) %>%
  sf::st_sf()
max(Torn.sf$yr)
```

Add energy dissipation per tornado.
```{r}
perc <- c(1, 0, 0, 0, 0, 0, 
          .772, .228, 0, 0, 0, 0,
          .616, .268, .115, 0, 0, 0,
          .529, .271, .133, .067, 0, 0,
          .543, .238, .131, .056, .032, 0,
          .538, .223, .119, .07, .033, .017)
percM <- matrix(perc, ncol = 6, byrow = TRUE)
threshW <- c(29.06, 38.45, 49.62, 60.8, 74.21, 89.41)
midptW <- c(diff(threshW)/2 + threshW[-length(threshW)], threshW[length(threshW)] + 7.5)
ef <- Torn.sf$mag + 1
EW3 <- numeric()
for(i in 1:length(ef)) EW3[i] = midptW^3 %*% percM[ef[i], ]
Torn.sf <- Torn.sf %>%
  mutate(ED = EW3 * AreaPath)
```

Add tornado paths by buffering tornado tracks by associated widths. Need to account for tornadoes without a path (same latitude and longitude).

```{r}
line <- Torn.sf %>%
  st_as_sf(coords = c("slon", "elon", "slat", "elat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("LINESTRING")
```

```{r}

TornL = as(Torn.sf, "Spatial")
TornP = gBuffer(TornL, byid = TRUE, width = TornL$Width/2, capStyle = "FLAT")
```



Plot casualty-producing tornadoes.
```{r}
sts <- state.name[!state.name %in% c("Alaska", "Hawaii")]
stateBorders <- us_states(states = sts)

tm_shape(stateBorders, projection ="+init=epsg:2163") +
  tm_borders() +
  tm_fill(col = "grey94") +
  tm_shape(Torn.sf$geometry) +
  tm_polygons(col = "black")
  tm_bubbles("Diff", col = "#fd8d3c", sizes.legend = c(50, 100, 500, 1000, 1500), alpha = 0.9, title.size="Underprediction", scale = 2.5) + 
  tm_shape(stateBorders) + 
  tm_borders(col = "grey", alpha = 0.3) +
  tm_format('World', legend.position = c("left", "bottom"),
                   attr.position = c("left", "bottom"),
                  legend.frame = FALSE) +
  #tm_format_Europe(legend.position = c("left", "bottom"),
  #                 attr.position = c("left", "bottom"),
  #                 legend.frame = TRUE) +
  tm_scale_bar(position = c("right", "bottom")) +
  #tm_compass(position = c("right", "bottom")) +
  tm_layout(frame = FALSE, attr.outside=TRUE)
```







